Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election

### 2024-03-12 paper/us_election_2020.qmd
last changes

---
title: "Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election"
author: 
  - Hannah Yu
  - Mary Cheng
  - Yimiao Yuan
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Election"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

raw_ces2020[1:5, 1:5] |>
  kable()
```

## Data Measurement

@@

The dataset is cleaned by rename the column names, specify the class of the columns, and change the numbers in the table to the corresponding description in the codebook to improve the readability. In addition, the cleaned data removes NA employment status since there are only 34 rows, which is a small amount of missing value compared with 61000 rows. After cleaning, 43534 rows of data and 4 variables remain.

The cleaned data will be analyzed and performed using R (@citeR) with `tidyverse` (@citeTidyverse), `here` (@citeHere), `rstanarm` (@citeRstanarm), `modelsummary` (@citeModleSummary), `ggplot2` (@citeGgplot2), `knitr` (@citeKnitr), `marginaleffects` (@citeMarginalEffects), `plotly` (@citePlotly), `tibble` (@citeTibble), `margins` (@citeMargins), and `kableExtra` (@citeKableExtra). @tbl-clean-data shows a preview of the cleaned dataset.

@@

@tbl-summary is a summary of the cleaned data, showing detailed statistics about the dataset. As we can see from the table, there are more people who support Biden. Respondents covered a wide range of races, with white people being the most heavily represented. Also, CES data came from a variety of locations, but the majority came from the southern region of the United States. The employment status of those who completed the survey varied, with most of them being full-time or retired.
```{r}
#| label: tbl-summary
#| tbl-cap: "Statistics summary of the cleaned 2020 CES dataset"
#| message: false
#| echo: false

@@

## Model set-up

@@

### Model justification

@@

# plot_ly(x=ces2020$race, y=ces2020$region, z=ces2020$employ, colors = c("red", "blue")) %>% 
#   add_markers(color = ces2020$voted_for, opacity = 0.03, size = 0.4) %>%
#   layout(xaxis = list(title = 'Race'), yaxis = list(title = 'Region'), zaxis = list(title = 'Employment Status')) %>%
#   layout(xaxis = list(titlefont = list(size = 22), tickfont = list(size = 22)),
#          yaxis = list(titlefont = list(size = 22), tickfont = list(size = 22)),
#          zaxis = list(titlefont = list(size = 22), tickfont = list(size = 22)))
```
This bubble plot (@fig-race_employ_region) illustrates four variables: race, region, employment status, and voting preference. Darker dots indicate a higher number of people, with red representing Trump supporters and blue for Biden supporters. Purple dots suggest similar support rates for both candidates, since blue and red makes purple in a 1:1 ratio. We can see that the darker dots are mostly in White, Black, Hispanic communities for full-time, part-time, or retired individuals. The darker dots are all blue, which indicates stronger support for Biden. It can also be noted that there are very few entirely red dots, which means that there is no significant disparities in voting percentages between Trump and Biden even among those groups which partially supported Trump. In those purple dots, we found that mostly are from people who are either Native American or from the South.

@@

Our results are summarized in @tbl-modelresults.

@@

modelsummary(
  list(
    "Support Biden" = political_preferences1
  ),
  statistic = "mad"
  )
```
Our results generally matches our justification. To avoid multicollinearity, the model excluded three variables from each category: race Asian, region Midwest, and employed full-time.The intercept represents the estimated log-odds of supporting Biden when all other predictors are held constant at their reference levels. In this case, the estimated log-odds of supporting Biden for individuals who are Asian, live in the Midwest, and are employed full-time is 0.806.

@@

Figure # shows range of coefficient estimates of our model within the 90% probability. Due to the fact that the credibility interval for race "Other" is particularly large, we cannot observe the trend of the 90% credibility intervals of other variables. Therefore we created figure # with the x axis limited from -10 to 10.

Combing figure # and figure #, we observe statistical significance for the coefficient estimates for students, the Northeast region, individuals identifying with two or more races, individuals identifying with "Other" racial category, Native Americans, Black individuals, and the intercept, Asians in the Midwest region who are employed full-time. The estimates are significant if the intervals do not cross 0. The value for the estimates are in log-odds, indicating that if the coefficient is positive, the individual supports Biden, if negative, the individual supports Trump.

@@

## First discussion point

In our analysis, respondents of different races show varying levels of support, with a majority of blacks, Hispanics, and Asians supporting Biden, and a majority of Middle Easterners and multiracials supporting Trump. The results for whites are more evenly distributed, with about the same percentage supporting Biden and Trump.

The reason for the majority of Hispanics supporting Biden may be due to the US-Mexico border wall that was built during Trump's reign. The wall was built over 452 miles, which replaced the old fence (@Rodriguez_2021). With this measure in place, Trump lost most of the Hispanic vote. Trump claims the wall reformed the immigration system and achieved the most secure southern border in US history. However, that wasn't true, and with it has come a surge in illegal border crossings. For Hispanics, Trump's long history of anti-immigrant rhetoric and this border wall have denied them the opportunity to seek asylum. While crossing the border illegally is not something to be promoted, it is the last hope for those living in dire straits.

2020 is the age of pandemics, as Trump's phrase "Chinese virus" led to a large number of Asians starting to support Biden. This is an extremely anti-Asian statement, although Covid was first found in China, but it does not mean that it is the Chinese virus. Trump's comments quickly festered in public opinion and led to a dramatic increase in hate incidents in the Asian American community (@Reja_2021). The repeated use of the "Chinese Virus" by a president of the country has contributed to an environment of racial discrimination and hatred against Asians, which has directly led to more Asians supporting Biden, whose policies are more open and democratic, and who values racial equality more.

Most white people vote relatively evenly, due to the fact that they are not affected by immigration or other policies. Whites hold a major voice and position in the US, most of the ideas and policies Trump proposes are favorable to whites. Therefore, he can gain support from most whites. However, there are some whites who want a democratic and free environment, and at the same time do not want to see racial discrimination and other phenomena, so they will choose to support Biden, who values equality.

@@

maybe talk about employment?
tax

@@

## Weaknesses and next steps

>>>>

---
title: "Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election"
author: 
  - Hannah Yu
  - Mary Cheng
  - Yimiao Yuan
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Election"
date: today
date-format: long
abstract: "This paper delves into the dynamics of voter turnout of the 2020 United States presidential election. Utilizing dataset from the Cooperate Election Study (CES2020), we employ logistic regression analysis to explore the influence of the voter's race, region, and employment on their vote for presidential candidate. Our findings reveal distinct voting patterns among different demographic groups, with minorities showing strong support for Biden, the South predominantly favored Trump, and individuals reliant on unemployment benefits preferred Biden. These insights stress the need for customized outreach and policies to enhance voter engagement across diverse U.S. communities."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

raw_ces2020[1:5, 1:5] |>
  kable()
```
## Data Measurement

@@

The dataset is cleaned by rename the column names, specify the class of the columns, and change the numbers in the table to the corresponding description in the codebook to improve the readability. In addition, the cleaned data removes NA employment status since there are only 34 rows, which is a small amount of missing value compared with 61000 rows. After []

The cleaned data will be analyzed and performed using R (@citeR) with `tidyverse` (@citeTidyverse), `here` (@citeHere), `rstanarm` (@citeRstanarm), `modelsummary` (@citeModleSummary), `ggplot2` (@citeGgplot2), `knitr` (@citeKnitr), `marginaleffects` (@citeMarginalEffects), `plotly` (@citePlotly), `tibble` (@citeTibble), `margins` (@citeMargins), `testthat` (@citetestthat) and `kableExtra` (@citeKableExtra). @tbl-clean-data shows a preview of the cleaned dataset.

@@

@tbl-summary is a summary of the cleaned data, showing detailed statistics about the dataset. ***add detail***
```{r}
#| label: tbl-summary
#| tbl-cap: "Statistics summary of the cleaned 2020 CES dataset"
#| message: false
#| echo: false

@@

## Model Set-up

@@

### Model Justification

@@

# plot_ly(x=ces2020$race, y=ces2020$region, z=ces2020$employ, colors = c("red", "blue")) %>% 
#   add_markers(color = ces2020$voted_for, opacity = 0.03, size = 0.4) %>%
#   layout(xaxis = list(title = 'Race'), yaxis = list(title = 'Region'), zaxis = list(title = 'Employment Status')) %>%
#   layout(xaxis = list(titlefont = list(size = 22), tickfont = list(size = 22)),
#          yaxis = list(titlefont = list(size = 22), tickfont = list(size = 22)),
#          zaxis = list(titlefont = list(size = 22), tickfont = list(size = 22)))
```

@@

@@

modelsummary(
  list(
    "Support Biden" = political_preferences1
  ),
  statistic = "mad"
  )
```
Our results are summarized in @tbl-modelresults1.Our results generally matches our justification. To avoid multicollinearity, the model excluded three variables from each category: race Asian, region Midwest, and employed full-time.The intercept represents the estimated log-odds of supporting Biden when all other predictors are held constant at their reference levels. In this case, the estimated log-odds of supporting Biden for individuals who are Asian, live in the Midwest, and are employed full-time is 0.806.

@@

@fig-modelresults1 shows range of coefficient estimates of our model within the 90% probability. Due to the fact that the credibility interval for race "Other" is particularly large, we cannot observe the trend of the 90% credibility intervals of other variables. Therefore we created figure # with the x axis limited from -10 to 10.

Combing @fig-modelresults1 and @fig-modelresults2, we observe statistical significance for the coefficient estimates for students, the Northeast region, individuals identifying with two or more races, individuals identifying with "Other" racial category, Native Americans, Black individuals, and the intercept, Asians in the Midwest region who are employed full-time. The estimates are significant if the intervals do not cross 0. The value for the estimates are in log-odds, indicating that if the coefficient is positive, the individual supports Biden, if negative, the individual supports Trump.

@@

## First discussion point
race
white
- religion

Biden: white, asian(trump covid), black, hispanic(trump建墙, no illegal immigration), other
Trump: white, two or more races, 
```{r}
ces2020|>
  filter(voted_for == "Biden") |>
  select(race) |>
  summary()
```
```{r}
ces2020|>
  filter(voted_for == "Trump") |>
  select(race) |>
  summary()
```

@@

maybe talk about employment?

@@

## Weaknesses and Next Steps
